Fundamental measure theory for hard polyhedra.

### 2013-01-31 papers/polyhedra/paper.tex
a few edits to polyhedron paper Tests skipped!

@@

We propose to treat the orientational dependence of the fundamental
measures using an expansion in spherical harmonics:
\begin{align}
  w_\alpha(\rr) &= \sum_{lm} \tilde{w}_\alpha^{lm}(|\rr|) Y_{lm}(\rhat)
\end{align}
Given this expansion, we can simplify the weighted densities:
\begin{align}
  n_\alpha(\rr) &= \int \sum_{lm} n(\rr',\rot)
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
  Y_{lm}(\widehat{R(\rot)(\rr-\rr')}) d\rr' d\rot
  \\
  &= \sum_{lmm'} \int \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
  Y_{lm'}(\widehat{(\rr-\rr')})
  n(\rr',\rot)
       D_{lmm'}(\rot) d\rot d\rr'
       \\
  &=
    \sum_{lmm'} \int \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
    Y_{lm'}(\widehat{(\rr-\rr')})
    n_{lmm'}(\rr') d\rr'
\end{align}
This reduces the challenge of handling the rotation to that of solving
for a set of densities $n_{lmm'}(\rr)$.
\subsection{Ideal gas free energy}
To address this, we will
first consider the free energy of an ideal gas.
\begin{align}
  A_{\textit{ideal}} &=
  k_BT \int
  \left(n(\rr,\rot)\ln(\Lambda^3\Theta^3 n(\rr,\rot)) - n(\rr,\rot)
  \right) d\rr d\rot
  \\
  &=
  k_BT \sum_{lmm'} \int
  n_{lmm'}(\rr)D_{lmm'}(\rot)\left(
  \ln\left(\Lambda^3\Theta^3 \sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)D_{l_1m_1m'_1}(\rot)\right) - 1
  \right) d\rr d\rot
  \\
  &=
  k_BT \int n_{000}(\rr)
  \left(\ln\left(\Lambda^3\Theta^3 \sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)\right) - 1
  \right) d\rr d\rot
  \\&+
  k_BT \sum_{l>1,mm'} \int
  n_{lmm'}(\rr)D_{lmm'}(\rot)
  \ln\left(\Lambda^3\Theta^3 \sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)D_{l_1m_1m'_1}(\rot)\right)
  d\rr d\rot
  \\
  &= A_{000} + \sum_{l>1,mm'} A_{lmm'} - k_BT\int n_{000}(\rr)d\rr
\end{align}
Here I have broken the $000$ term out of the outer summation over
$lmm'$, since this one is special, since it is independent of $\rot$.
We'll tackle that integral first, since it is a tad simpler.
\begin{align}
  A_{000} &= k_BT\int
  n_{000}(\rr)\ln\left(
  \sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)
  \right)
  d\rr d\rot
\end{align}
We can simplify this integral a bit if we write the log as a power
series...
\begin{align}
  A_{000}
  &= k_BT\int
  n_{000}(\rr)\ln\left(
  n_{000}(\rr)\left(1 + \frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  \right)
  d\rr d\rot
  \\
  &= k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)
  \\&+k_BT\int
  n_{000}(\rr)\ln\left(1 + \frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  d\rr d\rot
  \\
  &= k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)
  \\&+k_BT\int
  n_{000}(\rr)\left(\frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  d\rr d\rot
  \\&-\frac{k_BT}{2!}\int
  n_{000}(\rr)\left(\frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)^2
  d\rr d\rot
  \\&+\cdots
\end{align}
Now the linear term in the power series will vanish, since the
summation is over $l>0$ only.  In the quadratic term, only a few terms
will survive, and we'll end up with
\begin{align}
  A_{000}
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad-\frac{k_BT}{2!}\int
  \frac{\sum_{lmm'} \frac{2l+1}{8\pi^2} n_{lmm'}(\rr)n_{l,-m,-m'}}{n_{000}(\rr)}
  d\rr
  \\&\quad+\cdots
  \\
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad-\frac{k_BT}{2!}\int
  \frac{\sum_{lmm'} \frac{2l+1}{8\pi^2} |n_{lmm'}(\rr)|^2}{n_{000}(\rr)}
  d\rr
  \\&\quad+\cdots
\end{align}
Now an interesting question is whether we can mimic this power series
with a log, such as
\begin{align}
  A_{000}
  &=
  k_BT\int
  n_{000}(\rr)\ln \left( \sqrt{n_{000}(\rr)^2 + \sum_{lmm'}
    c_ln_{lmm'}(\rr)^2} \right)d\rr
  \\
  &= 
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad+
  k_BT\int n_{000}(\rr)\ln\left( \sqrt{1 + \frac{\sum_{lmm'}
    c_ln_{lmm'}(\rr)^2}{n_{000}(\rr)^2}} \right)d\rr
  \\
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad+
  k_BT\int \frac12 \frac{\sum_{lmm'}
    c_ln_{lmm'}(\rr)^2}{n_{000}(\rr)} d\rr
  \\&\quad+\cdots
\end{align}
from which we can conclude that to get an answer correct to the
quadratic order, we need
\begin{align}
  c_l &= -\frac{2l+1}{8\pi^2}
\end{align}
which seems all right except for the sign:  I expect that the free
energy should increase as we increase $n_{lmm'}$, since it means the
system is becoming more ordered, and it looks like to this quadratic
term, the free energy decreases... except that we're only looking at
$A_{000}$ right now, and thus are neglecting a possibly significant
increase due to the other terms...  so perhaps we should call this all
right, and move on for now.

@@

\subsection{Finding the orientational moments}
We find the orientational moments $n_{lmm'}(\rr)$ by minimizing the
free energy.  Let's begin by optimizing the free energy with respect
to nonzero angular momenta:
\begin{align}
  0&=\frac{\delta A}{\delta n_{lmm'}(\rr)} \\
  &= \frac{2l+1}{8\pi^2}n_{lmm'}(\rr)
  + k_BT \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \frac{\delta n_\alpha(\rr')}{\delta n_{lmm'}(\rr)} d\rr'
  \\
  &=
   \frac{2l+1}{8\pi^2}n_{lmm'}(\rr)
  \\&\quad+ k_BT \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)Y_{lm'}(\widehat{\rr-\rr'}) d\rr'
\end{align}
Solving for $n_{lmm'}(\rr)$, we find that
\begin{align}\label{eq:nlmm'}
  n_{lmm'}(\rr) &= \frac{8\pi^2k_BT}{2l+1}
  \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)Y_{lm'}(\widehat{\rr-\rr'}) d\rr'
\end{align}
I note that this equation for $n_{lmm'}$ is approximate, since we made
an approximation when we neglected higher order terms in the expansion
of the log in the ideal gas free energy.  Thus in cases where there is
relatively little orientational ordering, it ought to be an accurate
equation.  In any case I imagine it will be plausible.

@@

The final consideration (for $w_3^{000}(r)$ of a tetrahedron) is when
$R < r < \sqrt{3}R$, when we will want to just integrate over the
curvy triangles formed by the portion of the tetrahedron that stick
through the sphere.  We just need to integrate through one of these
points, and multiply by four... which might be easiest to do in a
Cartesian coordinates rotated by 45$^\circ$ around the ordinary $z$
axis.
\begin{align}
  w_3^{000}(r)
  &=
  \frac4{4\pi r^2}
  \int_{0}^R dz
  \int_0^{\frac{R+z}{\sqrt{2}}} dy
  \int_0^{\frac{R-z}{\sqrt{2}}} dx
  \delta(r-\sqrt{x^2+y^2+z^2})
  \\
  &=
  \frac1{\pi r^2}
  \int_{0}^R dz
  \int_0^{\frac{R+z}{\sqrt{2}}} dy
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \delta(r-\sqrt{x^2+y^2+z^2})
  \\
  u&= \sqrt{x^2+y^2+z^2}\\
  du &= \frac{ydy}{\sqrt{x^2+y^2+z^2}}
  \\
  w_3^{000}(r)
  &=
  \frac1{\pi r^2}
  \int_{0}^R dz
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \int_{\sqrt{z^2+x^2}}^{\sqrt{x^2+z^2+\frac{(R+z)^2}{2}}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2-x^2}}
  \\
  &=
  \frac1{\pi r^2}
  \int_{\sqrt{r^2-2R^2}}^R dz
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \int_{\sqrt{z^2+x^2}}^{\sqrt{x^2+z^2+\frac{(R+z)^2}{2}}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2-x^2}}
  \\
  z^2+x^2+y^2 &= r^2 \\
  z_{min}^2 + (R-z_{min})^2/2 &= r^2
\end{align}
Instead I'm going now to think of using cylindrical coordinates...
\begin{align}
  w_3^{000}(r) &= 3\frac{4}{4\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_0^{2\frac{\sqrt{3}R-z}{\cos\phi}} ds
  \delta(r-\sqrt{z^2+s^2})
  \\
  &= \frac{3}{\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_0^{2\frac{\sqrt{3}R-z}{\cos\phi}} ds
  \delta(r-\sqrt{z^2+s^2})
  \\
  u &= z^2+s^2 \\
  du &= sds \\
  &= \frac{\sqrt{u^2-z^2}}{u}ds
  \\
  w_3^{000}(r)
  &= \frac{3}{\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_{z}^{\sqrt{z^2+\left(2\frac{\sqrt{3}R-z}{\cos\phi}\right)^2}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2}}
  \\
  R &= \frac{2}{\sqrt{3}}s\cos\phi + \frac{1}{\sqrt{3}}z \\
  s &= 2\frac{\sqrt{3}R-z}{\cos\phi}
\end{align}
\newcommand\thetamax{\theta_{\text{max}}}
\begin{widetext}
Maybe spherical coordinates? Rotated $\pi/4$? Let's look first at the
plane that is our limit for $\theta$:
\begin{align}
  \sqrt{3}R &= x + y + z \\
  &= \sqrt{2}r\sin\thetamax\cos\phi +
  r\cos\thetamax
  \\
  \sqrt{3}\frac{R}{r} &=
  \sqrt{2}\sqrt{1-\cos^2\thetamax}\cos\phi
  + \cos\thetamax
  \\
  \sqrt{1-\cos^2\thetamax} &=
  \frac{\sqrt{3}\frac{R}{r}-\cos\thetamax}{\sqrt{2}\cos\phi}
  \\
  1-\cos^2\thetamax &=
  \frac{3\frac{R^2}{r^2}+\cos^2\thetamax -
    2\sqrt{3}\frac{R}{r}\cos\thetamax}{2\cos^2\phi}
  \\
  0 &=
  (1+2\cos^2\phi)\cos^2\thetamax
  - 2\sqrt{3}\frac{R}{r}\cos\thetamax
  + \left(3\frac{R^2}{r^2} - 2\cos^2\phi\right)
  \\
  \cos\thetamax &= \frac{2\sqrt{3}\frac{R}{r}
  - \sqrt{12\frac{R^2}{r^2} - 4(1+2\cos^2\phi)\left(3\frac{R^2}{r^2}-2\cos^2\phi\right)}}{2(1+2\cos^2\phi)}
  \\
  &= \frac{\sqrt{3}\frac{R}{r}
  - \sqrt{3\frac{R^2}{r^2} - (1+2\cos^2\phi)\left(3\frac{R^2}{r^2}-2\cos^2\phi\right)}}{1+2\cos^2\phi}
  \\
  &= \frac{\sqrt{3}\frac{R}{r}
  - \sqrt{3\frac{R^2}{r^2} - (2-\cos2\phi)\left(3\frac{R^2}{r^2}-1+\cos2\phi\right)}}{2-\cos2\phi}
\end{align}
\end{widetext}
This is sort of a mess, but does look useable.
\begin{align}
  w_3^{000}(r) &=
  \frac{12}{4\pi}
  \int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi
  \int_0^{\thetamax}d\theta
  \sin\theta
  \\
  &=
  \frac{3}{\pi}
  \int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}
  (1 - \cos\thetamax)d\phi
\end{align}
Surely we can either use Euler's relation or a geometrical
construction to solve for $\thetamax$?!

>>>>

\newcommand\thetamax{\theta_{\text{max}}}

@@

We propose to treat the orientational dependence of the fundamental
measures using an expansion in spherical harmonics:
\begin{align}
  w_\alpha(\rr) &= \sum_{lm} \tilde{w}_\alpha^{lm}(|\rr|) Y_{lm}(\rhat)
\end{align}
Given this expansion, we can simplify the weighted densities:
\begin{align}
  n_\alpha(\rr) &= \int \sum_{lm} n(\rr',\rot)
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
  Y_{lm}(\widehat{R(\rot)(\rr-\rr')}) d\rr' d\rot
  \\
  &= \sum_{lmm'} \int \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
  Y_{lm'}(\widehat{(\rr-\rr')})
  n(\rr',\rot)
       D_{lmm'}(\rot) d\rot d\rr'
       \\
  &=
    \sum_{lmm'} \int \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
    Y_{lm'}(\widehat{(\rr-\rr')})
    n_{lmm'}(\rr') d\rr'
\end{align}
This reduces the challenge of handling the rotation to that of solving
for a set of densities $n_{lmm'}(\rr)$.

\subsection{Ideal gas free energy}
To address this, we will
first consider the free energy of an ideal gas.
\begin{widetext}
\begin{align}
  A_{\textit{ideal}} &=
  k_BT \int
  \left(n(\rr,\rot)\ln(\Lambda^3\Theta^3 n(\rr,\rot)) - n(\rr,\rot)
  \right) d\rr d\rot
  \\
  &=
  k_BT \sum_{lmm'} \int
  n_{lmm'}(\rr)D_{lmm'}(\rot)\left(
  \ln\left(\Lambda^3\Theta^3 \sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)D_{l_1m_1m'_1}(\rot)\right) - 1
  \right) d\rr d\rot
  \\
  &=
  k_BT \int n_{000}(\rr)
  \left(\ln\left(\Lambda^3\Theta^3 \sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)\right) - 1
  \right) d\rr d\rot
  \\&+
  k_BT \sum_{l>1,mm'} \int
  n_{lmm'}(\rr)D_{lmm'}(\rot)
  \ln\left(\Lambda^3\Theta^3 \sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)D_{l_1m_1m'_1}(\rot)\right)
  d\rr d\rot
  \\
  &= A_{000} + \sum_{l>1,mm'} A_{lmm'} - k_BT\int n_{000}(\rr)d\rr
\end{align}
Here I have broken the $000$ term out of the outer summation over
$lmm'$, since this one is special, since it is independent of $\rot$.
We'll tackle that integral first, since it is a tad simpler.
\begin{align}
  A_{000} &= k_BT\int
  n_{000}(\rr)\ln\left(
  \sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)
  \right)
  d\rr d\rot
\end{align}
We can simplify this integral a bit if we write the log as a power
series...
\begin{align}
  A_{000}
  &= k_BT\int
  n_{000}(\rr)\ln\left(
  n_{000}(\rr)\left(1 + \frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  \right)
  d\rr d\rot
  \\
  &= k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)
  \\&+k_BT\int
  n_{000}(\rr)\ln\left(1 + \frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  d\rr d\rot
  \\
  &= k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)
  \\&+k_BT\int
  n_{000}(\rr)\left(\frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  d\rr d\rot
  \\&-\frac{k_BT}{2!}\int
  n_{000}(\rr)\left(\frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)^2
  d\rr d\rot
  \\&+\cdots
\end{align}
Now the linear term in the power series will vanish, since the
summation is over $l>0$ only.  In the quadratic term, only a few terms
will survive, and we'll end up with
\begin{align}
  A_{000}
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad-\frac{k_BT}{2!}\int
  \frac{\sum_{lmm'} \frac{2l+1}{8\pi^2} n_{lmm'}(\rr)n_{l,-m,-m'}}{n_{000}(\rr)}
  d\rr
  \\&\quad+\cdots
  \\
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad-\frac{k_BT}{2!}\int
  \frac{\sum_{lmm'} \frac{2l+1}{8\pi^2} |n_{lmm'}(\rr)|^2}{n_{000}(\rr)}
  d\rr
  \\&\quad+\cdots
\end{align}
Now an interesting question is whether we can mimic this power series
with a log, such as
\begin{align}
  A_{000}
  &=
  k_BT\int
  n_{000}(\rr)\ln \left( \sqrt{n_{000}(\rr)^2 + \sum_{lmm'}
    c_ln_{lmm'}(\rr)^2} \right)d\rr
  \\
  &= 
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad+
  k_BT\int n_{000}(\rr)\ln\left( \sqrt{1 + \frac{\sum_{lmm'}
    c_ln_{lmm'}(\rr)^2}{n_{000}(\rr)^2}} \right)d\rr
  \\
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad+
  k_BT\int \frac12 \frac{\sum_{lmm'}
    c_ln_{lmm'}(\rr)^2}{n_{000}(\rr)} d\rr
  \\&\quad+\cdots
\end{align}
from which we can conclude that to get an answer correct to the
quadratic order, we need
\begin{align}
  c_l &= -\frac{2l+1}{8\pi^2}
\end{align}
which seems all right except for the sign:  I expect that the free
energy should increase as we increase $n_{lmm'}$, since it means the
system is becoming more ordered, and it looks like to this quadratic
term, the free energy decreases... except that we're only looking at
$A_{000}$ right now, and thus are neglecting a possibly significant
increase due to the other terms...  so perhaps we should call this all
right, and move on for now.

@@

\end{widetext}
\subsection{Finding the orientational moments}
We find the orientational moments $n_{lmm'}(\rr)$ by minimizing the
free energy.  Let's begin by optimizing the free energy with respect
to nonzero angular momenta:
\begin{align}
  0&=\frac{\delta A}{\delta n_{lmm'}(\rr)} \\
  &= \frac{2l+1}{8\pi^2}n_{lmm'}(\rr)
  + k_BT \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \frac{\delta n_\alpha(\rr')}{\delta n_{lmm'}(\rr)} d\rr'
  \\
  &=
   \frac{2l+1}{8\pi^2}n_{lmm'}(\rr)
  \\&\quad+ k_BT \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)Y_{lm'}(\widehat{\rr-\rr'}) d\rr'
\end{align}
Solving for $n_{lmm'}(\rr)$, we find that
\begin{align}\label{eq:nlmm'}
  n_{lmm'}(\rr) &= \frac{8\pi^2k_BT}{2l+1}
  \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)Y_{lm'}(\widehat{\rr-\rr'}) d\rr'
\end{align}
I note that this equation for $n_{lmm'}$ is approximate, since we made
an approximation when we neglected higher order terms in the expansion
of the log in the ideal gas free energy.  Thus in cases where there is
relatively little orientational ordering, it ought to be an accurate
equation.  In any case I imagine it will be plausible.

@@

The final consideration (for $w_3^{000}(r)$ of a tetrahedron) is when
$R < r < \sqrt{3}R$, when we will want to just integrate over the
curvy triangles formed by the portion of the tetrahedron that stick
through the sphere.  We just need to integrate through one of these
points, and multiply by four... which might be easiest to do in a
Cartesian coordinates rotated by 45$^\circ$ around the ordinary $z$
axis.
\begin{widetext}
\begin{align}
  w_3^{000}(r)
  &=
  \frac4{4\pi r^2}
  \int_{0}^R dz
  \int_0^{\frac{R+z}{\sqrt{2}}} dy
  \int_0^{\frac{R-z}{\sqrt{2}}} dx
  \delta(r-\sqrt{x^2+y^2+z^2})
  \\
  &=
  \frac1{\pi r^2}
  \int_{0}^R dz
  \int_0^{\frac{R+z}{\sqrt{2}}} dy
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \delta(r-\sqrt{x^2+y^2+z^2})
  \\
  u&= \sqrt{x^2+y^2+z^2}\\
  du &= \frac{ydy}{\sqrt{x^2+y^2+z^2}}
  \\
  w_3^{000}(r)
  &=
  \frac1{\pi r^2}
  \int_{0}^R dz
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \int_{\sqrt{z^2+x^2}}^{\sqrt{x^2+z^2+\frac{(R+z)^2}{2}}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2-x^2}}
  \\
  &=
  \frac1{\pi r^2}
  \int_{\sqrt{r^2-2R^2}}^R dz
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \int_{\sqrt{z^2+x^2}}^{\sqrt{x^2+z^2+\frac{(R+z)^2}{2}}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2-x^2}}
  \\
  z^2+x^2+y^2 &= r^2 \\
  z_{min}^2 + (R-z_{min})^2/2 &= r^2
\end{align}
Instead I'm going now to think of using cylindrical coordinates...
\begin{align}
  w_3^{000}(r) &= 3\frac{4}{4\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_0^{2\frac{\sqrt{3}R-z}{\cos\phi}} ds
  \delta(r-\sqrt{z^2+s^2})
  \\
  &= \frac{3}{\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_0^{2\frac{\sqrt{3}R-z}{\cos\phi}} ds
  \delta(r-\sqrt{z^2+s^2})
  \\
  u &= z^2+s^2 \\
  du &= sds \\
  &= \frac{\sqrt{u^2-z^2}}{u}ds
  \\
  w_3^{000}(r)
  &= \frac{3}{\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_{z}^{\sqrt{z^2+\left(2\frac{\sqrt{3}R-z}{\cos\phi}\right)^2}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2}}
  \\
  R &= \frac{2}{\sqrt{3}}s\cos\phi + \frac{1}{\sqrt{3}}z \\
  s &= 2\frac{\sqrt{3}R-z}{\cos\phi}
\end{align}
Maybe spherical coordinates? Rotated $\pi/4$? Let's look first at the
plane that is our limit for $\theta$:
\begin{align}
  \sqrt{3}R &= x + y + z \\
  &= \sqrt{2}r\sin\thetamax\cos\phi +
  r\cos\thetamax
  \\
  \sqrt{3}\frac{R}{r} &=
  \sqrt{2}\sqrt{1-\cos^2\thetamax}\cos\phi
  + \cos\thetamax
  \\
  \sqrt{1-\cos^2\thetamax} &=
  \frac{\sqrt{3}\frac{R}{r}-\cos\thetamax}{\sqrt{2}\cos\phi}
  \\
  1-\cos^2\thetamax &=
  \frac{3\frac{R^2}{r^2}+\cos^2\thetamax -
    2\sqrt{3}\frac{R}{r}\cos\thetamax}{2\cos^2\phi}
  \\
  0 &=
  (1+2\cos^2\phi)\cos^2\thetamax
  - 2\sqrt{3}\frac{R}{r}\cos\thetamax
  + \left(3\frac{R^2}{r^2} - 2\cos^2\phi\right)
  \\
  \cos\thetamax &= \frac{2\sqrt{3}\frac{R}{r}
  - \sqrt{12\frac{R^2}{r^2} - 4(1+2\cos^2\phi)\left(3\frac{R^2}{r^2}-2\cos^2\phi\right)}}{2(1+2\cos^2\phi)}
  \\
  &= \frac{\sqrt{3}\frac{R}{r}
  - \sqrt{3\frac{R^2}{r^2} -
    (1+2\cos^2\phi)\left(3\frac{R^2}{r^2}-2\cos^2\phi\right)}}{1+2\cos^2\phi}
  \\
  &= \frac{\sqrt{3}\frac{R}{r}
  - \sqrt{3\frac{R^2}{r^2} - (2-\cos2\phi)\left(3\frac{R^2}{r^2}-1+\cos2\phi\right)}}{2-\cos2\phi}
\end{align}
\end{widetext}
This is sort of a mess, but does look useable.
\begin{align}
  w_3^{000}(r) &=
  \frac{12}{4\pi}
  \int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi
  \int_0^{\thetamax}d\theta
  \sin\theta
  \\
  &=
  \frac{3}{\pi}
  \int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}
  (1 - \cos\thetamax)d\phi
\end{align}
Surely we can either use Euler's relation or a geometrical
construction to solve for $\thetamax$?!
